\documentclass{letter}

% my address and name
\address{University of Arizona\\
         Dept. of Chemistry\\
         1306 E. University\\
         Tucson, AZ 85721\\
         \medskip
         phone: (520)621-6761\\
         e-mail: \texttt{kinghorn@u.arizona.edu}\\
         Fax: (520) 621-6761}

\signature{Dr. Donald B. Kinghorn}

\begin{document}

% the recipient
\begin{letter}{Prof. E. R. Davidson, Associate Editor, The Journal of Chemical Physics\\
               Dept. of Chemistry,\\
               Indiana University\\
               Bloomington, IN 47405}      

% the opening greeting
\opening{Dear Prof. Davidson}
Regarding the manuscript, \textbf{A correlated basis set for
non-adiabatic energy calculations on diatomic molecules }
by Donald B. Kinghorn and Ludwik Adamowicz. 

A8.11.151

We have revised the manuscript as suggested by the referee and are returning
two copies of the revised manuscript along with the transfer of copyright
agreement per your request.

The following revisions have been made:
\begin{quote}
  Page 6: added the parenthetical statement --- 
  Also, we use a back tick, ('), to denote the matrix/vector transpose.---\\
  Page 7: added the parenthetical statement --- The Kronecker product, or 
  matrix outer product, is defined in the references$^{6}$ or any
  advanced linear algebra text.---\\
  Page 16: set $\mathbf{a_{j}}$ to math bold in text and eqn. (28)\\
  Page 17: deleted text --- ... and the transpose of A is denoted A'---
\end{quote}

We have the inclosed manuscript in LaTeX format and can submit it electronically if this is convenient.

\encl{Revised manuscript (2 copies ), transfer of copyright agreement}

\closing{Warmest Regards}

\end{letter}

\end{document}
