\documentstyle[12pt,overcite]{article}
\setlength{\textwidth}{7in}
\setlength{\textheight}{9.25in}
\setlength{\oddsidemargin}{-.25in}
%
\setlength{\topmargin}{-.9in}
\def\Real{\rm I\hspace{-0.2em}R}
\def\nin{{\in \! \! \! \! \! /} \,}
\def\nequiv{{\equiv \! \! \! \! \! \! /} \,}



\begin{document}
\setcounter{page}{32}

\begin{center}
{\Large 
\bf 
Budget Justification}
\end{center}
 
\vspace{2mm}

\setlength{\baselineskip}{2em}

We have two main budget needs in this project. First, we
need to provide summer salary support for the P.I.
and salaries 
for  the postdoctoral research associate
and 
for the graduate student
working on the project.
The proposed budget will provide two months of salary for the
P.I. and twelve months of salary for a 
postdoctoral research associate and a graduate student,
each year.


The second budget item is for the purchase of
% an SGI ORIGIN 200
%4-cpu system with 1GB RAM and 20 GB disk. 
20  nodes for the construction of our new parallel Beowulf cluster.
Which will be based on the 4 processor prototype cluster we
are currently building.
Each pair of nodes
will be housed in a dual processor ``box'' of high end 
PC components including 2 Pentium II 400 MHz processors and 
256MB of 8ns 100MHz memory. Communication will be over 
switched 100Mb Fast Ethernet. Through our own experience
and conversations with others who have built similar clusters,
we expect performance at least as good as a 48 processor
wide node SP2 super computer for our applications. Building this 
cluster  utilizing commodity PC components gives orders of
magnitude increases in performance/cost over conventional
multiprocessor workstations.
This system 
will provide an excellent parallel environment for the 
proposed method development and application calculations.
%At the price of \$30,554 this system is probably the best
%buy at the present time. The SGI computer will complement
%and eventually replace
%our aging cluster of IBM RISK 6000 workstations.

The third budget need are the travel expenses and
operation funds for the project, including materials
and supplies, publication costs, computer services,
communication costs, {\it etc.}
We budgeted the minimum what will be needed for this 
purpose.


\end{document}



