%% This document created by Scientific Word (R) Version 2.0


\documentclass[12pt,thmsa]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{sw20jart}

%TCIDATA{TCIstyle=Article/art4.lat,jart,sw20jart}

\input tcilatex
\QQQ{Language}{
American English
}

\begin{document}

After two or three years of postdoctoral training I will be seeking an
academic position at a research University. My plans are to establish a
vigorous research group capable of making important advances in theoretical
and computational quantum chemistry. In addition to the research and
development activities of this group I would like to provide a computational
service to other research groups, making the powerful tools of computational
quantum chemistry available to a wide range of experimental and theoretical
researchers. I would actively seek collaboration with other researchers in
chemistry, physics, biology, mathematics and computer science.

In addition to my research activities I am very interested in education. I
find teaching to be both very enjoyable and challenging. My choice of an
academic career will allow the pursuit of  both my research and teaching
interests. I expect to be active in undergraduate and graduate education.

Also, I have plans to write a monograph on matrix differential calculus. My
recent publications are the first in the scientific literature to utilize
this powerful new mathematics. I am currently preparing lecture notes on
this subject for a series of seminars I am presenting to the group I am
working with here in Arizona. I will probably be able to start work on the
manuscript in three of four years.

I consider postdoctoral training an essential element in my career plans. It
affords time to broaden my experience as a theoretical chemist and to
establish a history of quality research. An NSF fellowship would greatly
increase the probability of my securing a favorable position in the highly
competitive academic job market.

\end{document}
