%% This document created by Scientific Word (R) Version 2.0
%% Starting shell: mathart1


\documentclass[12pt,thmsa]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\usepackage{sw20aip}
\usepackage{doublesp}

%TCIDATA{Created=Mon Apr 15 10:38:32 1996}
%TCIDATA{LastRevised=Wed Jun 05 14:18:19 1996}
%TCIDATA{Language=American English}

\input tcilatex
\begin{document}

\author{Donald B. Kinghorn and Ludwik Adamowicz \\
%EndAName
Deptartment of Chemistry, \\
University of Arizona \\
Tucson, AZ 85721, U.S.A.}
\title{The Electron Affinity of Hydrogen, Deuterium, and Tritium: A Non-adiabatic
Variational Calculation Using Explicitly Correlated Gaussian Basis Functions}
\date{June 4, 1996}
\maketitle

\begin{abstract}
Improved values for the electron affinities of Hydrogen, Deuterium and
Tritium are reported. Values are 6083.9168$cm^{-1}$, 6086.7137$cm^{-1}$, and
6087.9168$cm^{-1}$ for Hydrogen, Deuterium and Tritium, respectively. These
results were obtained from non-adiabatic variational energy calculations in
a basis of correlated Gaussians. Energy bounds from these calculations
surpass previously reported results and new ``exact'' nonrelativistic energy
bounds are carefully predicted: $E\left( H^{-}\right)
=-.5274458811,\,\,E\left( D^{-}\right) =-.5275983247$ and $E\left(
T^{-}\right) =-.5276490482$ in hartree energy units. It is shown that these
new bounds cannot be obtained using the infinite nuclear mass approximation
plus Rydberg scaling and the ``exact'' mass polarization correction.
\end{abstract}

\section{Introduction}

Experimental values for the electron affinity of Hydrogen have been steadily
improving. In the mid 70's a value of 6081$\pm 16cm^{-1}$ was reported by
Feldmann\cite{Feldmann75} and independently by McCulloh and Walker\cite
{McCulloh74}. Greater accuracy was then achieved by Chupka, Dehmer and Jivery%
\cite{Chupka75} giving 6083$\left( +11,-3\right) cm^{-1}$. In 1979 Scherk%
\cite{Scherk79} reported a value of 6085.5$\pm 3.3cm^{-1}$. Then in 1991
Lykke, Murray and Lineberger\cite{Lykke91} achieved a 20 fold improvement in
accuracy obtaining 6082.99$\pm 0.15cm^{-1}$ and 6086.2$\pm 0.6cm^{-1}$ for
the electron affinity of Hydrogen and Deuterium respectively. Recent
conversation with Professor Lineberger suggested that improved error limits
especially for Deuterium may be possible. In anticipation of these improved
experimental results  the present work gives improved theoretical
predictions for the electron affinity of Hydrogen, Deuterium and Tritium. 

There have been several highly accurate ground state energy calculations for
the negative Hydrogen ion\cite
{Pekeris62,Frankowski66,Drake88,Baker90,Thakkar94,Ackermann95}. The most
accurate of these calculations\cite{Drake88,Baker90,Thakkar94} appear to be
converged to the point where the uncertainties in the calculated energies is
less than $10^{-13}$ hartree ($10^{-8}cm^{-1}$). This uncertainty is several
orders of magnitude less than the uncertainty in even the most accurately
known fundamental physical constants. For example, the Rydberg constant is
uncertain in the 10th digit\cite{Petley92} and the atomic mass of Hydrogen
is uncertain in the 11th digit\cite{Audi93} (the electron mass is uncertain
in the 8th digit\cite{Audi93}). As remarkable as these calculations are they
suffer from one major shortcoming ---the computed wave functions are for a
fictitious system with an infinitely heavy nuclear mass. It is common to add
corrections to these computed energies. The corrections consist of scaling
using a reduced-mass-corrected Rydberg constant, the reduced Rydberg, and
adding what is usually called a mass polarization correction\cite
{BetheSalpeter}. These corrections to the infinite nuclear mass
approximation are very good, however, they do not account for the entire
energy shift due to nuclear motion even when ``exact'' mass polarization
corrections are used. For highly accurate energy calculations on the systems 
$H^{-},\,\,D^{-},\,\,$and $T^{-}$ (and probably for Helium) it is necessary
to use a Hamiltonian that explicitly includes the nuclear motion ---the
non-adiabatic Hamiltonian. We demonstrate this in the present work by
providing new rigorous upper bounds for the non-relativistic ground state
energy of $H^{-}$ and $D^{-}$ (our $T^{-}$ result is not quite a new bound).
These new bounds are computed directly using the finite mass non-adiabatic
Hamiltonian. Our $H^{-}$ energy bound is nearly 5 nano-hartree lower than
can be achieved using the infinite nuclear mass approximation plus Rydberg
scaling and even exact mass polarization corrections, and we predict that
the exact energy lowering should be closer to 9 nano-hartrees. It appears
that this 9 nano-hartrees can only be recovered rigorously using the
non-adiabatic Hamiltonian. This is an energy correction approximately half
that of the Lamb shift (17 nano-hartrees\cite{Drake88}) and thus represents
a significant deficiency in the infinite nuclear mass approximation. We use
our new non-relativistic energy bounds together with relativistic and other
small corrections computed by Drake\cite{Drake88} to obtain what should be
reliable theoretical values for the electron affinity of Hydrogen, Deuterium
and Tritium. We hope that this work will spark renewed interest in high
accuracy electron affinity determination among both experimentalists and
theoreticians.

\section{Theory}

\subsection{The Hamiltonian}

In general the total Hamiltonian for a three-particle system with masses $%
\left\{ M_1,M_2,M_3\right\} $ and charges $\left\{ Q_1,Q_2,Q_3\right\} $
interacting under a Coulomb potential is represented (in atomic units) by
\begin{equation}
H_{tot}=\frac{P_1^2}{2M_1}+\frac{P_2^2}{2M_2}+\frac{P_3^2}{2M_3}+\frac{Q_1Q_2%
}{R_{12}}+\frac{Q_1Q_3}{R_{13}}+\frac{Q_2Q_3}{R_{23}}  \label{Htot}
\end{equation}
Where $R_i$, $P_i$ are position and momentum vectors for particle $i$ and $%
R_{ij}=\left\| R_i-R_j\right\| .$ This Hamiltonian is separable into a
Hamiltonian for the kinetic energy of the center of mass and a Hamiltonian
describing the internal interaction energy. A transformation leading to this
separation is given by\cite{Poshusta83}
\begin{equation}
\begin{array}{llll}
r_0=\frac{M_1R_1+M_2R_2+M_3R_3}{m_0}, & p_0=P_1+P_2+P_3, & m_0=M_1+M_2+M_{3,}
& q_0=Q_1 \\ 
r_1=R_2-R_1, & p_1=P_2-\frac{M_2}{m_0}p_0, & m_1=\frac{M_1M_2}{M_1+M_2}, & 
q_1=Q_2 \\ 
r_2=R_3-R_1, & p_2=P_3-\frac{M_3}{m_0}p_0, & m_2=\frac{M_2M_3}{M_2+M_3}, & 
q_2=Q_3
\end{array}
\label{CMtran}
\end{equation}
Applying this transformation to the total Hamiltonian gives $H_{tot}=H_{cm}+H
$ where
\begin{equation}
H_{cm}=\frac{p_0^2}{2m_0}
\end{equation}
and
\begin{equation}
H=\frac{p_1^2}{2m_1}+\frac{p_2^2}{2m_2}+\frac{p_1\cdot p_2}{M_1}+\frac{q_0q_1%
}{r_1}+\frac{q_0q_2}{r_2}+\frac{q_1q_2}{r_{12}}  \label{H1}
\end{equation}
The Hamiltonian in eqn.(\ref{H1}) can be viewed as representing the internal
motions of a three-particle system or as the total energy of a two-particle
system with fictitious masses $m_1$, $m_2$ and charges $q_1,\,q_2$
interacting with a charge $q_0$ at the origin together with their mutual
Coulomb interaction and a momentum dependent ``mass polarization potential''.

Now, for $H^{-}$, $D^{-}$, and $T^{-}$, we assign the nucleus to particle
one and the electrons to particles two and three giving the nonrelativistic,
nonadiabatic, Hamiltonian used in these calculations 
\begin{equation}
H=-\frac{\nabla _1^2}{2\mu }-\frac{\nabla _2^2}{2\mu }-\frac{\nabla _1\cdot
\nabla _2}{M_N}-\frac 1{r_1}-\frac 1{r_2}+\frac 1{r_{12}}
\end{equation}
Here $\mu $ is the reduced electron mass, $M_N$ is the nuclear mass, $r_1$
and $r_2$ are relative coordinates described in the center of mass
transformation, eqn.(\ref{CMtran}), and we have written the momentum
operator $p$ as $-i\nabla $. For the values of the nuclear masses see table(%
\ref{masstab}).

\subsection{Wave Function}

Calculations where performed using explicitly correlated Gaussian basis
functions of the form ( $^{\prime }$ represents vector/matrix transposition
and $\otimes $ is the Kronecker product symbol) 
\begin{eqnarray}
\phi _k &=&\exp \left[ -r^{\prime }\left( L_kL_k^{\prime }\otimes I_3\right)
r\right]   \label{ECG} \\
&=&\exp \left[ -r^{\prime }\left( A_k\otimes I_3\right) r\right]   \nonumber
\\
&=&\exp \left[ -\sum_{i,j}a_{ij}\left( r_i\cdot r_j\right) \right]  
\nonumber
\end{eqnarray}
Where, in general, for an $n$ particle (pseudo-particle) system $r$ is the $%
3n\times 1$ vector of Cartesian coordinates for the $n$ particles, $L_k$ is
an $n\times n$ rank $n$ lower triangular matrix of nonlinear variation
parameters, and $I_3$ is the $3\times 3$ identity matrix. The Kronecker
product with the identity insures rotational invariance of the basis
functions. The exponent parameters are written in Cholesky factored form, $%
L_kL_k^{\prime },$ to insure positive definiteness of the quadratic form. 
Correlation is more obvious when one notes the relation $r_{ij}^2=r_i\cdot
r_i+r_j\cdot r_j-2r_i\cdot r_j.$ This form of the basis is equivalent to the
(single center) form of the basis introduced by Boys\cite{Boys} and then, in
matrix form, by Singer\cite{Singer1} in 1960. For a more complete discussion
of this basis and derivation of the Hamiltonian matrix elements and
derivatives in matrix form see Kinghorn\cite{Kinghorn95a,Kinghorn95b}. For
the particular cases we are considering in this work $r=\left( r_1^{\prime
},\,r_2^{\prime }\right) ^{\prime }$ is the $6\times 1$ vector of relative
coordinates defined above. The ground state spatial wave function (symmetric
with respect to exchange of electrons) is then given as the symmetry
projected linear combination of the $\phi _k$, 
\begin{equation}
\Psi =\sum_kc_k\left( \exp \left[ -r^{\prime }\left( L_kL_k^{\prime }\otimes
I_3\right) r\right] +\exp \left[ -r^{\prime }\left( \tau ^{\prime
}L_kL_k^{\prime }\tau \otimes I_3\right) r\right] \right) 
\end{equation}
where $\tau $ is the permutation matrix $\left( 
\begin{array}{cc}
0 & 1 \\ 
1 & 0
\end{array}
\right) .$  

\subsection{Optimization}

The wave functions are optimized for the ground state energy by minimizing
the Rayleigh quotient 
\begin{equation}
E\left( a;c\right) =\min_{\left\{ a,c\right\} }\frac{c^{\prime }H(a)c}{%
c^{\prime }S(a)c}  \label{energy}
\end{equation}
$H\left( a\right) $ and $S\left( a\right) $ are the Hamiltonian and overlap
matrices which are functions of the nonlinear parameters contained in the
basis set exponent matrices $L_k$. We write $a$ for the collection of these
nonlinear parameters. $c$ is the vector of linear coefficients in the basis
expansion of $\Psi .$ Thus, if we let $N$ be the number of basis functions,
the energy is, in general, a function of $Nn\left( n+1\right) /2+N$
variables which in our particular case is $4N$ variables. Obviously a
reduction of $N$ variables could be achieved by solving the eigen-problem
associated with eqn. (\ref{energy}) to obtain the linear coefficients and
then iteratively optimize the nonlinear parameters. However, we found that
much more thorough optimization could be achieved by letting the optimizer
simultaneously vary both the linear and nonlinear parameters.

The optimization software employed was the package TN by Stephen Nash\cite
{NashTN} --- available from netlib\cite{netlib}. TN is a truncated Newton
method utilizing a user supplied gradient. The analytic gradient of the
energy functional was derived using matrix differential calculus\cite
{Kinghorn95a,Kinghorn95b}.

\subsection{Mass values and other constants}

The values for nuclear masses and other constants used in these calculations
are presented in table(\ref{masstab}). The nuclear masses where computed
using the atomic masses given in \emph{The 1993 atomic mass evaluation} of
Audi and Wapstra\cite{Audi93}. The nuclear mass is derived from the atomic
mass using 
\begin{equation}
M_N=M_A-Z\times m_e+B_e
\end{equation}
Here $M_A$ is the atomic mass, $Z$ the nuclear charge, $m_e$ the electron
mass, and $B_e$ the total electron binding energy. $B_e$ is the negative of
the ground state energy and is converted to mass units using $%
E=mc^2=m/\alpha ^2.$ We use quantum units in this work except where
otherwise noted. Thus, $\hbar =1,$ $m_e=1,$ energy is in hartree$\left(
=2R_\infty \right) $, and distance in bohr. The value of the Rydberg
constant, $R_\infty $, was taken from CODATA86\cite{codata86} and the fine
structure constant, $\alpha $, from Kinoshita\cite{Kinoshita95}. The finite
nuclear mass Hydrogen like atom ground state energy differs from the
infinite nuclear mass result by the factor $\mu /m_e=1-\mu /M_N=M_N/\left(
M_N+m_e\right) .$ This factor is used later in this work when effect of
nuclear motion are discussed.

It should be noted that the electron mass contains error in its eighth
significant digit ($m_e=548579.903\pm 0.013$ nano-atomic units)\cite{Audi93}%
, this, together with the error in the nuclear masses, propagates to an
error of approximately $\pm 6\times 10^{-10}$ in the reduced masses $\mu _i$%
. Thus, even the ``exact'' ground state energy for Hydrogen is in error in
the 10th digit. Therefore, to reflect (at least) the uncertainties in the
fundamental physical constants, results in this work will be presented with
at most 10 digits.

%TCIMACRO{\TeXButton{B}{\begin{table}[tbp] \centering}}
%BeginExpansion
\begin{table}[tbp] \centering%
%EndExpansion
\begin{tabular}{ll}
\hline\hline
$M_H=1836.152693$ & $\mu _H=.9994556794$ \\ 
$M_D=3670.483008$ & $\mu _D=.9997276305$ \\ 
$M_T=5496.921571$ & $\mu _T=.9998181131$ \\ \hline
$R_\infty =109737.31534cm^{-1}$ &  \\ 
$\alpha ^{-1}=137.03599944$ &  \\ \hline
$E_0=-\mu /2$ &  \\ 
$E_0\left( H\right) =-.4997278397$ &  \\ 
$E_0\left( D\right) =-.4998638152$ &  \\ 
$E_0\left( T\right) =-.4999090565$ &  \\ \hline\hline
\end{tabular}
\caption{Mass values and other constants \label{masstab}}%
%TCIMACRO{\TeXButton{E}{\end{table}}}
%BeginExpansion
\end{table}%
%EndExpansion

\section{Results}

\subsection{Expectation values}

Tables (\ref{Hinfexpvals}, \ref{Hexpvals},\ref{Dexpvals},\ref{Texpvals})
contain expectation values for Hydrogen (infinite nuclear mass
approximation), Hydrogen, Deuterium , and Tritium respectively. Results
using wave functions with from 64 to 512 terms are included to exhibit
convergence which we feel is surprisingly good for this basis. Included in
the tables are expectation values for the Hamiltonian , the kinetic and
potential energy, the virial coefficient ($\eta $), components of the
kinetic and potential energy, integrals for $r,$ $r^2,$ and $r^{-2}$ and
finally contact integrals.

%TCIMACRO{\TeXButton{B}{\begin{table}[tbp] \centering} }
%BeginExpansion
\begin{table}[tbp] \centering%
%EndExpansion
\begin{tabular}{lllll}
\hline\hline
$N$ & $64$ & $128$ & $256$ & $512$ \\ \hline
$\left\langle H\right\rangle _\infty $ & \multicolumn{1}{r}{-.5277504812} & 
\multicolumn{1}{r}{-.5277509852} & \multicolumn{1}{r}{-.5277510093} & 
\multicolumn{1}{r}{-.5277510116} \\ 
$\left\langle T\right\rangle _\infty $ & \multicolumn{1}{r}{.5277504436} & 
\multicolumn{1}{r}{.5277509709} & \multicolumn{1}{r}{.5277510703} & 
\multicolumn{1}{r}{.5277510096} \\ 
$\left\langle V\right\rangle _\infty $ & \multicolumn{1}{r}{-1.0555009248} & 
\multicolumn{1}{r}{-1.0555019562} & \multicolumn{1}{r}{-1.0555020796} & 
\multicolumn{1}{r}{-1.0555020212} \\ 
$\eta $ & \multicolumn{1}{r}{1.0000000355} & \multicolumn{1}{r}{1.0000000135}
& \multicolumn{1}{r}{.9999999422} & \multicolumn{1}{r}{1.0000000019} \\ 
$\left\langle \nabla _1^2\right\rangle _\infty $ & \multicolumn{1}{r}{
.5277504436} & \multicolumn{1}{r}{.5277509709} & \multicolumn{1}{r}{
.5277510702} & \multicolumn{1}{r}{.5277510095} \\ 
$\left\langle \nabla _1\cdot \nabla _2\right\rangle _\infty $ & 
\multicolumn{1}{r}{.0328798325} & \multicolumn{1}{r}{.0328797865} & 
\multicolumn{1}{r}{.0328797911} & \multicolumn{1}{r}{.0328797798} \\ 
$\left\langle r_1^{-1}\right\rangle _\infty $ & \multicolumn{1}{r}{
.6832622892} & \multicolumn{1}{r}{.6832617853} & \multicolumn{1}{r}{
.6832618093} & \multicolumn{1}{r}{.6832617605} \\ 
$\left\langle r_{12}^{-1}\right\rangle _\infty $ & \multicolumn{1}{r}{
.3110236535} & \multicolumn{1}{r}{.3110216145} & \multicolumn{1}{r}{
.3110215390} & \multicolumn{1}{r}{.3110214997} \\ 
$\left\langle r_1\right\rangle _\infty $ & \multicolumn{1}{r}{2.7101080515}
& \multicolumn{1}{r}{2.7101737546} & \multicolumn{1}{r}{2.7101774512} & 
\multicolumn{1}{r}{2.7101782272} \\ 
$\left\langle r_{12}\right\rangle _\infty $ & \multicolumn{1}{r}{4.4125559666
} & \multicolumn{1}{r}{4.4126857296} & \multicolumn{1}{r}{4.4126929713} & 
\multicolumn{1}{r}{4.4126944006} \\ 
$\left\langle r_1^2\right\rangle _\infty $ & \multicolumn{1}{r}{11.9117467188
} & \multicolumn{1}{r}{11.9135603852} & \multicolumn{1}{r}{11.9136812624} & 
\multicolumn{1}{r}{11.9136972111} \\ 
$\left\langle r_{12}^2\right\rangle _\infty $ & \multicolumn{1}{r}{
25.1981384317} & \multicolumn{1}{r}{25.2017491899} & \multicolumn{1}{r}{
25.2019887841} & \multicolumn{1}{r}{25.2020204446} \\ 
$\left\langle r_1^{-2}\right\rangle _\infty $ & \multicolumn{1}{r}{
1.1166299163} & \multicolumn{1}{r}{1.1166554041} & \multicolumn{1}{r}{
1.1166570728} & \multicolumn{1}{r}{1.1166583435} \\ 
$\left\langle r_{12}^{-2}\right\rangle _\infty $ & \multicolumn{1}{r}{
.1551216229} & \multicolumn{1}{r}{.1551059984} & \multicolumn{1}{r}{
.1551054860} & \multicolumn{1}{r}{.1551054766} \\ 
$\left\langle \delta \left( r_1\right) \right\rangle _\infty $ & 
\multicolumn{1}{r}{.1635839048} & \multicolumn{1}{r}{.1640737909} & 
\multicolumn{1}{r}{.1640744018} & \multicolumn{1}{r}{.1641346963} \\ 
$\left\langle \delta \left( r_{12}\right) \right\rangle _\infty $ & 
\multicolumn{1}{r}{.0028090290} & \multicolumn{1}{r}{.0027610047} & 
\multicolumn{1}{r}{.0027593265} & \multicolumn{1}{r}{.0027596593} \\ 
\hline\hline
\end{tabular}
\caption{$^{\infty}$H$^{-}$ Expectation Values \label{Hinfexpvals}}%
%TCIMACRO{\TeXButton{E}{\end{table}}}
%BeginExpansion
\end{table}%
%EndExpansion

%TCIMACRO{\TeXButton{B}{\begin{table}[tbp] \centering} }
%BeginExpansion
\begin{table}[tbp] \centering%
%EndExpansion
\begin{tabular}{lllll}
\hline\hline
$N$ & $64$ & $128$ & $256$ & $512$ \\ \hline
$\left\langle H\right\rangle _H$ & \multicolumn{1}{r}{-.5274453611} & 
\multicolumn{1}{r}{-.5274458498} & \multicolumn{1}{r}{-.5274458739} & 
\multicolumn{1}{r}{-.5274458762} \\ 
$\left\langle T\right\rangle _H$ & \multicolumn{1}{r}{.5274453148} & 
\multicolumn{1}{r}{.5274458206} & \multicolumn{1}{r}{.5274458565} & 
\multicolumn{1}{r}{.5274458636} \\ 
$\left\langle V\right\rangle _H$ & \multicolumn{1}{r}{-1.0548906759} & 
\multicolumn{1}{r}{-1.0548916704} & \multicolumn{1}{r}{-1.0548917304} & 
\multicolumn{1}{r}{-1.0548917398} \\ 
$\eta $ & \multicolumn{1}{r}{1.0000000438} & \multicolumn{1}{r}{1.0000000276}
& \multicolumn{1}{r}{1.0000000165} & \multicolumn{1}{r}{1.0000000119} \\ 
$\left\langle \nabla _1^2\right\rangle _H$ & \multicolumn{1}{r}{.5271403731}
& \multicolumn{1}{r}{.5271408787} & \multicolumn{1}{r}{.5271409145} & 
\multicolumn{1}{r}{.5271409216} \\ 
$\left\langle \nabla _1\cdot \nabla _2\right\rangle _H$ & \multicolumn{1}{r}{
.0327790347} & \multicolumn{1}{r}{.0327789973} & \multicolumn{1}{r}{
.0327790013} & \multicolumn{1}{r}{.0327790010} \\ 
$\left\langle r_1^{-1}\right\rangle _H$ & \multicolumn{1}{r}{.6828538727} & 
\multicolumn{1}{r}{.6828533936} & \multicolumn{1}{r}{.6828533723} & 
\multicolumn{1}{r}{.6828533730} \\ 
$\left\langle r_{12}^{-1}\right\rangle _H$ & \multicolumn{1}{r}{.3108170695}
& \multicolumn{1}{r}{.3108151168} & \multicolumn{1}{r}{.3108150142} & 
\multicolumn{1}{r}{.3108150062} \\ 
$\left\langle r_1\right\rangle _H$ & \multicolumn{1}{r}{2.7120310559} & 
\multicolumn{1}{r}{2.7120911055} & \multicolumn{1}{r}{2.7120952238} & 
\multicolumn{1}{r}{2.7120955275} \\ 
$\left\langle r_{12}\right\rangle _H$ & \multicolumn{1}{r}{4.4155651280} & 
\multicolumn{1}{r}{4.4156837795} & \multicolumn{1}{r}{4.4156918099} & 
\multicolumn{1}{r}{4.4156924114} \\ 
$\left\langle r_1^2\right\rangle _H$ & \multicolumn{1}{r}{11.9300411389} & 
\multicolumn{1}{r}{11.9316074492} & \multicolumn{1}{r}{11.9317345507} & 
\multicolumn{1}{r}{11.9317438071} \\ 
$\left\langle r_{12}^2\right\rangle _H$ & \multicolumn{1}{r}{25.2337827345}
& \multicolumn{1}{r}{25.2368966127} & \multicolumn{1}{r}{25.2371493544} & 
\multicolumn{1}{r}{25.2371678863} \\ 
$\left\langle r_1^{-2}\right\rangle _H$ & \multicolumn{1}{r}{1.1153654889} & 
\multicolumn{1}{r}{1.1153908649} & \multicolumn{1}{r}{1.1153923541} & 
\multicolumn{1}{r}{1.1153937757} \\ 
$\left\langle r_{12}^{-2}\right\rangle _H$ & \multicolumn{1}{r}{.1549104913}
& \multicolumn{1}{r}{.1548959719} & \multicolumn{1}{r}{.1548954385} & 
\multicolumn{1}{r}{.1548954516} \\ 
$\left\langle \delta \left( r_1\right) \right\rangle _H$ & 
\multicolumn{1}{r}{.1633128785} & \multicolumn{1}{r}{.1638020720} & 
\multicolumn{1}{r}{.1638024979} & \multicolumn{1}{r}{.1638623602} \\ 
$\left\langle \delta \left( r_{12}\right) \right\rangle _H$ & 
\multicolumn{1}{r}{.0027992322} & \multicolumn{1}{r}{.0027545364} & 
\multicolumn{1}{r}{.0027528569} & \multicolumn{1}{r}{.0027531889} \\ 
\hline\hline
\end{tabular}
\caption{H$^{-}$ Expectation Values \label{Hexpvals}}%
%TCIMACRO{\TeXButton{E}{\end{table}}}
%BeginExpansion
\end{table}%
%EndExpansion

%TCIMACRO{\TeXButton{B}{\begin{table}[tbp] \centering} }
%BeginExpansion
\begin{table}[tbp] \centering%
%EndExpansion
\begin{tabular}{lllll}
\hline\hline
$N$ & $64$ & $128$ & $256$ & $512$ \\ \hline
$\left\langle H\right\rangle _D$ & \multicolumn{1}{r}{-.5275977894} & 
\multicolumn{1}{r}{-.5275982934} & \multicolumn{1}{r}{-.5275983175} & 
\multicolumn{1}{r}{-.5275983198} \\ 
$\left\langle T\right\rangle _D$ & \multicolumn{1}{r}{.5275977498} & 
\multicolumn{1}{r}{.5275982638} & \multicolumn{1}{r}{.5275983000} & 
\multicolumn{1}{r}{.5275983071} \\ 
$\left\langle V\right\rangle _D$ & \multicolumn{1}{r}{-1.0551955392} & 
\multicolumn{1}{r}{-1.0551965573} & \multicolumn{1}{r}{-1.0551966175} & 
\multicolumn{1}{r}{-1.0551966269} \\ 
$\eta $ & \multicolumn{1}{r}{1.0000000375} & \multicolumn{1}{r}{1.0000000279}
& \multicolumn{1}{r}{1.0000000164} & \multicolumn{1}{r}{1.0000000119} \\ 
$\left\langle \nabla _1^2\right\rangle _D$ & \multicolumn{1}{r}{.5274451065}
& \multicolumn{1}{r}{.5274456205} & \multicolumn{1}{r}{.5274456566} & 
\multicolumn{1}{r}{.5274456637} \\ 
$\left\langle \nabla _1\cdot \nabla _2\right\rangle _D$ & \multicolumn{1}{r}{
.0328293594} & \multicolumn{1}{r}{.0328293211} & \multicolumn{1}{r}{
.0328293246} & \multicolumn{1}{r}{.0328293249} \\ 
$\left\langle r_1^{-1}\right\rangle _D$ & \multicolumn{1}{r}{.6830579287} & 
\multicolumn{1}{r}{.6830574141} & \multicolumn{1}{r}{.6830573931} & 
\multicolumn{1}{r}{.6830573937} \\ 
$\left\langle r_{12}^{-1}\right\rangle _D$ & \multicolumn{1}{r}{.3109203181}
& \multicolumn{1}{r}{.3109182709} & \multicolumn{1}{r}{.3109181688} & 
\multicolumn{1}{r}{.3109181606} \\ 
$\left\langle r_1\right\rangle _D$ & \multicolumn{1}{r}{2.7110671655} & 
\multicolumn{1}{r}{2.7111329939} & \multicolumn{1}{r}{2.7111370994} & 
\multicolumn{1}{r}{2.7111374073} \\ 
$\left\langle r_{12}\right\rangle _D$ & \multicolumn{1}{r}{4.4140556422} & 
\multicolumn{1}{r}{4.4141856237} & \multicolumn{1}{r}{4.4141936235} & 
\multicolumn{1}{r}{4.4141942386} \\ 
$\left\langle r_1^2\right\rangle _D$ & \multicolumn{1}{r}{11.9207707499} & 
\multicolumn{1}{r}{11.9225874044} & \multicolumn{1}{r}{11.9227141243} & 
\multicolumn{1}{r}{11.9227234416} \\ 
$\left\langle r_{12}^2\right\rangle _D$ & \multicolumn{1}{r}{25.2157135447}
& \multicolumn{1}{r}{25.2193293841} & \multicolumn{1}{r}{25.2195812302} & 
\multicolumn{1}{r}{25.2196000783} \\ 
$\left\langle r_1^{-2}\right\rangle _D$ & \multicolumn{1}{r}{1.1159970928} & 
\multicolumn{1}{r}{1.1160225252} & \multicolumn{1}{r}{1.1160240187} & 
\multicolumn{1}{r}{1.1160254418} \\ 
$\left\langle r_{12}^{-2}\right\rangle _D$ & \multicolumn{1}{r}{.1550164901}
& \multicolumn{1}{r}{.1550008728} & \multicolumn{1}{r}{.1550003393} & 
\multicolumn{1}{r}{.1550003524} \\ 
$\left\langle \delta \left( r_1\right) \right\rangle _D$ & 
\multicolumn{1}{r}{.1634484158} & \multicolumn{1}{r}{.1639378072} & 
\multicolumn{1}{r}{.1639383038} & \multicolumn{1}{r}{.1639982438} \\ 
$\left\langle \delta \left( r_{12}\right) \right\rangle _D$ & 
\multicolumn{1}{r}{.0028057413} & \multicolumn{1}{r}{.0027577672} & 
\multicolumn{1}{r}{.0027560874} & \multicolumn{1}{r}{.0027564205} \\ 
\hline\hline
\end{tabular}
\caption{D$^{-}$ Expectation Values \label{Dexpvals}}%
%TCIMACRO{\TeXButton{E}{\end{table}}}
%BeginExpansion
\end{table}%
%EndExpansion

%TCIMACRO{\TeXButton{B}{\begin{table}[tbp] \centering} }
%BeginExpansion
\begin{table}[tbp] \centering%
%EndExpansion
\begin{tabular}{lllll}
\hline\hline
$N$ & $64$ & $128$ & $256$ & $512$ \\ \hline
$\left\langle H\right\rangle _T$ & \multicolumn{1}{r}{-.5276485136} & 
\multicolumn{1}{r}{-.5276490169} & \multicolumn{1}{r}{-.5276490410} & 
\multicolumn{1}{r}{-.5276490433} \\ 
$\left\langle T\right\rangle _T$ & \multicolumn{1}{r}{.5276484962} & 
\multicolumn{1}{r}{.5276489885} & \multicolumn{1}{r}{.5276490236} & 
\multicolumn{1}{r}{.5276490306} \\ 
$\left\langle V\right\rangle _T$ & \multicolumn{1}{r}{-1.0552970098} & 
\multicolumn{1}{r}{-1.0552980055} & \multicolumn{1}{r}{-1.0552980646} & 
\multicolumn{1}{r}{-1.0552980739} \\ 
$\eta $ & \multicolumn{1}{r}{1.0000000165} & \multicolumn{1}{r}{1.0000000268}
& \multicolumn{1}{r}{1.0000000164} & \multicolumn{1}{r}{1.0000000120} \\ 
$\left\langle \nabla _1^2\right\rangle _T$ & \multicolumn{1}{r}{.5275465495}
& \multicolumn{1}{r}{.5275470418} & \multicolumn{1}{r}{.5275470769} & 
\multicolumn{1}{r}{.5275470838} \\ 
$\left\langle \nabla _1\cdot \nabla _2\right\rangle _T$ & \multicolumn{1}{r}{
.0328461179} & \multicolumn{1}{r}{.0328460647} & \multicolumn{1}{r}{
.0328460796} & \multicolumn{1}{r}{.0328460808} \\ 
$\left\langle r_1^{-1}\right\rangle _T$ & \multicolumn{1}{r}{.6831258386} & 
\multicolumn{1}{r}{.6831253023} & \multicolumn{1}{r}{.6831252803} & 
\multicolumn{1}{r}{.6831252808} \\ 
$\left\langle r_{12}^{-1}\right\rangle _T$ & \multicolumn{1}{r}{.3109546673}
& \multicolumn{1}{r}{.3109525990} & \multicolumn{1}{r}{.3109524960} & 
\multicolumn{1}{r}{.3109524877} \\ 
$\left\langle r_1\right\rangle _T$ & \multicolumn{1}{r}{2.7107482436} & 
\multicolumn{1}{r}{2.7108143219} & \multicolumn{1}{r}{2.7108183967} & 
\multicolumn{1}{r}{2.7108187065} \\ 
$\left\langle r_{12}\right\rangle _T$ & \multicolumn{1}{r}{4.4135568905} & 
\multicolumn{1}{r}{4.4136872973} & \multicolumn{1}{r}{4.4136952742} & 
\multicolumn{1}{r}{4.4136958928} \\ 
$\left\langle r_1^2\right\rangle _T$ & \multicolumn{1}{r}{11.9177685588} & 
\multicolumn{1}{r}{11.9195898850} & \multicolumn{1}{r}{11.9197143110} & 
\multicolumn{1}{r}{11.9197236656} \\ 
$\left\langle r_{12}^2\right\rangle _T$ & \multicolumn{1}{r}{25.2098664126}
& \multicolumn{1}{r}{25.2134904549} & \multicolumn{1}{r}{25.2137388014} & 
\multicolumn{1}{r}{25.2137576949} \\ 
$\left\langle r_1^{-2}\right\rangle _T$ & \multicolumn{1}{r}{1.1162073447} & 
\multicolumn{1}{r}{1.1162327408} & \multicolumn{1}{r}{1.1162342331} & 
\multicolumn{1}{r}{1.1162356567} \\ 
$\left\langle r_{12}^{-2}\right\rangle _T$ & \multicolumn{1}{r}{.1550507158}
& \multicolumn{1}{r}{.1550357895} & \multicolumn{1}{r}{.1550352550} & 
\multicolumn{1}{r}{.1550352681} \\ 
$\left\langle \delta \left( r_1\right) \right\rangle _T$ & 
\multicolumn{1}{r}{.1634933913} & \multicolumn{1}{r}{.1639829810} & 
\multicolumn{1}{r}{.1639835043} & \multicolumn{1}{r}{.1640434722} \\ 
$\left\langle \delta \left( r_{12}\right) \right\rangle _T$ & 
\multicolumn{1}{r}{.0028046832} & \multicolumn{1}{r}{.0027588443} & 
\multicolumn{1}{r}{.0027571630} & \multicolumn{1}{r}{.0027574964} \\ 
\hline\hline
\end{tabular}
\caption{T$^{-}$ Expectation Values \label{Texpvals}}%
%TCIMACRO{\TeXButton{E}{\end{table}}}
%BeginExpansion
\end{table}%
%EndExpansion

The ground state energy of $H^{-}$ in the infinite nuclear mass
approximation has been computed to incredible precision with the best
results appearing to be converged to at least 13 digits. However, the
infinite nuclear mass approximation introduces errors in expectation values
in the 4th digit as is easily seen by comparing results in table (\ref
{Hinfexpvals}) with table (\ref{Hexpvals}). The standard corrections to the
energy introduced by Bethe and Salpeter\cite{BetheSalpeter}, ``reduced
Rydberg scaling'' and mass polarization corrections, leave error in the 8th
digit as will be demonstrated in section(\ref{nucmotian}). Even the Rydberg
constant, $R_\infty $, which would be used for energy unit conversion is
uncertain in the 10th digit\cite{Petley92}. With this clearly in mind we
include table (\ref{Litdata}) listing expectation values from other infinite
nuclear mass calculations. We were unable to find any rigorous nonadiabatic
calculations for $H^{-}$ in the literature.

%TCIMACRO{\TeXButton{B}{\begin{table}[tbp] \centering}}
%BeginExpansion
\begin{table}[tbp] \centering%
%EndExpansion
\begin{tabular}{lll}
\hline\hline
$\left\langle H\right\rangle _\infty $ & -.5277510062954 & {\footnotesize %
Pekeris}\cite{Pekeris62}{\footnotesize , 444 term Hylleraas type} \\ 
& -.52775101635 & {\footnotesize Frankowski and Pekeris}\cite{Frankowski66}%
{\footnotesize , 246 term Hylleraas + }$\ln $ \\ 
& -.527751016544203 & {\footnotesize Drake}\cite{Drake88}{\footnotesize ,
616 term Hylleraas with double exponents} \\ 
& -.527751016544375 & {\footnotesize Baker et al }\cite{Baker90}%
{\footnotesize , 476 term Hylleraas + }$\ln ,${\footnotesize \ 1/Z expansion}
\\ 
& -.527751016544240 & {\footnotesize Thakkar and Koga}\cite{Thakkar94}%
{\footnotesize , 455 term modified Hylleraas type} \\ 
& -.527751016532 & {\footnotesize Ackermann}\cite{Ackermann95}{\footnotesize %
, 2005 point Finite Element Method} \\ 
& -.5277510116 & {\footnotesize This work, 512 term correlated Gaussian} \\ 
\hline
$\left\langle \nabla _1\cdot \nabla _2\right\rangle _\infty $ & .0328797771
& {\footnotesize Pekeris}\cite{Pekeris62} \\ 
& .03287978125 & {\footnotesize Drake}\cite{Drake88} \\ 
& .0328797798 & {\footnotesize This work} \\ \hline
$\left\langle \delta \left( r_1\right) \right\rangle _\infty $ & .164547396
& {\footnotesize Pekeris}\cite{Pekeris62} \\ 
& .164552873 & {\footnotesize Drake}\cite{Drake88} \\ 
& .1641346963 & {\footnotesize This work} \\ \hline
$\left\langle \delta \left( r_{12}\right) \right\rangle _\infty $ & 
.002740206 & {\footnotesize Pekeris}\cite{Pekeris62} \\ 
& .002737995 & {\footnotesize Drake}\cite{Drake88} \\ 
& .002759659 & {\footnotesize This work} \\ \hline
$\left\langle r_1\right\rangle _\infty $ & 2.71017831 & {\footnotesize %
Pekeris}\cite{Pekeris62} \\ 
& 2.710178263 & {\footnotesize Ackermann}\cite{Ackermann95} \\ 
& 2.710178227 & {\footnotesize This work} \\ \hline
$\left\langle r_{12}\right\rangle _\infty $ & 4.41269452 & {\footnotesize %
Pekeris}\cite{Pekeris62} \\ 
& 4.412694401 & {\footnotesize This work} \\ \hline
$\left\langle r_1^2\right\rangle _\infty $ & 11.913692 & {\footnotesize %
Pekeris}\cite{Pekeris62} \\ 
& 11.913699235 & {\footnotesize Ackermann}\cite{Ackermann95} \\ 
& 11.9136972111 & {\footnotesize This work} \\ \hline
$\left\langle r_{12}^2\right\rangle _\infty $ & 25.202010 & {\footnotesize %
Pekeris}\cite{Pekeris62} \\ 
& 25.20202044 & {\footnotesize This work} \\ \hline\hline
\end{tabular}
\caption{$H^{-}$  infinite nuclear mass literature values \label{Litdata}}%
%TCIMACRO{\TeXButton{E}{\end{table}}}
%BeginExpansion
\end{table}%
%EndExpansion

\subsection{Finite Nuclear Mass Effects\label{nucmotian}\label{NMeffect}}

The energy shift due to finite nuclear mass is 
\begin{equation}
\Delta E_{NM}=\left\langle H\right\rangle _f-\left\langle H\right\rangle
_\infty   \label{dENM}
\end{equation}
Following Bethe and Salpeter\cite{BetheSalpeter} $\Delta E_{NM}$ can be
given to very good approximation, for two electron systems, by 
\begin{eqnarray}
\widetilde{\Delta E_{NM}} &=&\left[ \left( \frac{M_N}{m_e+M_N}\right)
\left\langle H\right\rangle _\infty +\left( \frac{M_N}{m_e+M_N}\right) \frac{%
m_e}{M_N}\left\langle \nabla _1\cdot \nabla _2\right\rangle _\infty \right]
-\left\langle H\right\rangle _\infty   \nonumber \\
&=&\left[ \left( 1-\frac \mu {M_N}\right) \left\langle H\right\rangle
_\infty +\frac \mu {M_N}\left\langle \nabla _1\cdot \nabla _2\right\rangle
_\infty \right] -\left\langle H\right\rangle _\infty   \nonumber \\
&=&-\frac \mu {M_N}\left( \left\langle H\right\rangle _\infty -\left\langle
\nabla _1\cdot \nabla _2\right\rangle _\infty \right)   \label{dNMapprox}
\end{eqnarray}
The first term in square brackets in eqn.(\ref{dNMapprox}) is the ``reduced
Rydberg scaling'' which could be interpreted as the energy change that would
result from changing units from hartree$(=2R_\infty )$ to reduced Rydberg
units $2R_N=2(\mu /m_e)R_\infty $. The second term is a perturbation, the
``mass polarization energy'', again in reduced Rydberg units.

In table (\ref{NMtab}) values for $\Delta E_{NM}$, $\widetilde{\Delta E_{NM}}
$ and $\Delta =\widetilde{\Delta E_{NM}}-\Delta E_{NM}$ are given using our
infinite nuclear mass wavefunction and non-adiabatic wave functions for
Hydrogen, Deuterium and Tritium anions.

%TCIMACRO{\TeXButton{B}{\begin{table}[tbp] \centering}}
%BeginExpansion
\begin{table}[tbp] \centering%
%EndExpansion
\begin{tabular}{lll}
\hline\hline
$\Delta E_{NM}\left( H^{-}\right) $ & .0003051354 & 66.96948$cm^{-1}$ \\ 
$\widetilde{\Delta E_{NM}}\left( H^{-}\right) $ & .0003051629 & 
\multicolumn{1}{r}{66.97551$cm^{-1}$} \\ 
$\Delta \left( H^{-}\right) $ & $2.75\times 10^{-8}$ & \multicolumn{1}{r}{
.00603$cm^{-1}$} \\ \hline
$\Delta E_{NM}\left( D^{-}\right) $ & .0001526918 & \multicolumn{1}{r}{
33.51198$cm^{-1}$} \\ 
$\widetilde{\Delta E_{NM}}\left( D^{-}\right) $ & .0001526987 & 
\multicolumn{1}{r}{33.51350$cm^{-1}$} \\ 
$\Delta \left( D^{-}\right) $ & $6.89\times 10^{-9}$ & \multicolumn{1}{r}{
.00151$cm^{-1}$} \\ \hline
$\Delta E_{NM}\left( T^{-}\right) $ & .0001019683 & \multicolumn{1}{r}{
22.37946$cm^{-1}$} \\ 
$\widetilde{\Delta E_{NM}}\left( T^{-}\right) $ & .0001019714 & 
\multicolumn{1}{r}{22.38014$cm^{-1}$} \\ 
$\Delta \left( T^{-}\right) $ & $3.08\times 10^{-9}$ & \multicolumn{1}{r}{
.00068$cm^{-1}$} \\ \hline\hline
\end{tabular}
\caption{Energy shift due to finite nuclear mass\label{NMtab}}%
%TCIMACRO{\TeXButton{E}{\end{table}}}
%BeginExpansion
\end{table}%
%EndExpansion

The approximation to the nuclear mass energy shift for $H^{-}$ leaves 27
nano-hartree, .00603$cm^{-1}$, unaccounted for which is unacceptable for
high accuracy calculations. Drake\cite{Drake88} has computed a total mass
polarization correction given as 
\begin{equation}
\Delta E_{M_H}\left( \text{Drake}\right) =.03287978125\left( \mu
_H/M_H\right) -.059779493\left( \mu _H/M_H\right) ^2.  \label{drakeEM}
\end{equation}
The first order term in this correction gives a shift of 3.927968$cm^{-1}$
which agrees with our calculation for $\frac{\mu _H}{M_H}\left\langle \nabla
_1\cdot \nabla _2\right\rangle _\infty $. The second order term however
gives only .003887$cm^{-1}$ of the .00603$cm^{-1}$ deficiency from the
approximation $\widetilde{\Delta E_{NM}}\left( H^{-}\right) $ of eqn.(\ref
{dNMapprox}). The coefficient of the first term in eqn.(\ref{drakeEM}) is $%
\left\langle \nabla _1\cdot \nabla _2\right\rangle _\infty $. The second
term is derived by subtracting the leading term in eqn.(\ref{drakeEM}) from
the total energy shift due to mass polarization. This total mass
polarization energy shift is computed by subtracting the infinite nuclear
mass energy from the energy derived from the Hamiltonian 
\begin{equation}
H=\frac 12(\nabla _1^2+\nabla _2^2)+\frac \mu M\left\langle \nabla _1\cdot
\nabla _2\right\rangle +V
\end{equation}
We followed this procedure and obtained 
\begin{equation}
\Delta E_{M_H}\left( \text{This work}\right) =.0328797798\left( \mu
_H/M_H\right) -.059834351\left( \mu _H/M_H\right) ^2,
\end{equation}
in reasonable agreement with Drake. Again, the second term gives only .003891%
$cm^{-1}$ of the .00603$cm^{-1}$ deficiency from the approximation $%
\widetilde{\Delta E_{NM}}\left( H^{-}\right) .$ Continuing this analysis for
Deuterium and Tritium gives
\begin{equation}
\Delta E_{M_D}\left( \text{This work}\right) =.0328797798\left( \mu
_D/M_D\right) -.059983567\left( \mu _D/M_D\right) ^2.
\end{equation}
Here 2.4 nano-hartree (.00053$cm^{-1}$) of the energy shift due to nuclear
motion is unaccounted for. For Tritium
\begin{equation}
\Delta E_{M_T}\left( \text{This work}\right) =.0328797798\left( \mu
_T/M_T\right) -.059983567\left( \mu _T/M_T\right) ^2
\end{equation}
leaves 1.9 nano-hartree (.00042$cm^{-1}$) unaccounted for. We must conclude
that Rydberg scaling and mass polarization alone can not account entirely
for the energy shift due to nuclear motion. 

This section is concluded with table(\ref{boundstab}) presenting the
variational energy upper bounds computed in this work. For comparison we
list the bounds derivable from Drake's work. Drake's work was chosen for two
reasons; 1) the infinite nuclear mass energy calculations are excellent and
2) data was reported to allow exact mass polarization corrections for $H^{-}$%
, and very good first order corrections for $D^{-}$ and $T^{-}$. Table(\ref
{boundstab}) includes our estimation of the exact energy upper bounds. Our
infinite nuclear mass energy value lies 4.9 nano-hartrees above the
essentially exact value reported by Drake and we believe that our
non-adiabatic calculations contain this same deficiency. Therefore, we
estimate the exact energy upper bounds by correcting our computed values for
this deficiency.

Note: In table 3 of Drake's 1988 paper\cite{Drake88} Rydberg scaling was
inadvertently applied to the first and second order mass polarization
corrections which were already scaled by nature of the way they were
computed ( eqn.(\ref{drakeEM}) in this paper ). This is an easily overlooked
error but fortunately easily corrected.

%TCIMACRO{\TeXButton{B}{\begin{table}[tbp] \centering}}
%BeginExpansion
\begin{table}[tbp] \centering%
%EndExpansion
\begin{tabular}{llll}
\hline\hline
& $H^{-}$ & $D^{-}$ & $T^{-}$ \\ \hline
Drake\cite{Drake88} & -.5274458714 & -.5275983178 & -.5276490451 \\ 
This work & -.5274458762 & -.5275983198 & -.5276490433 \\ 
``Exact'' estimate & -.5274458811 & -.5275983247 & -.5276490482 \\ 
\hline\hline
\end{tabular}
\caption{Energy upper bounds \label{boundstab}}%
%TCIMACRO{\TeXButton{E}{\end{table}}}
%BeginExpansion
\end{table}%
%EndExpansion

\subsection{Electron Affinity}

Electron affinities for Hydrogen Deuterium and Tritium are presented in
table (\ref{EAtab}). In the calculation of the electron affinities the
energies for the neutral atoms are those given in table(\ref{masstab}). The
calculations include highly accurate small corrections computed by Drake\cite
{Drake88}. These include relativistic, relativistic recoil, Lamb shift and
finite nuclear size corrections giving a total correction of 0.307505$cm^{-1}
$,  labeled  $\Delta E_{corr}$ in table (\ref{EAtab}). We use this
correction with Rydberg scaling for Deuterium and Tritium. The electron
affinity is then given by 
\begin{equation}
EA=E_{\text{atom}}-E_{\text{anion}}-\Delta E_{corr}
\end{equation}
where $E_{\text{atom}}$ is the energy of the neutral atom, $E_{\text{anion}}$
is the non-adiabatic energy of the negative ion and $\Delta E_{corr}$ is the
correction described above. Included in table(\ref{EAtab}) are the
computational results for Hydrogen of Pekeris\cite{Pekeris62} and of Drake%
\cite{Drake88}, and the experimental results for Hydrogen and Deuterium of
Lykke, Murray and Lineberger\cite{Lykke91}. Also included is our estimates
of the exact electron affinity computed from the exact energy estimates
given in table(\ref{boundstab}).

The result from Drake's work listed in our table(\ref{EAtab}) is the
corrected result (see the note at the end of section(\ref{NMeffect})). It is
interesting to note that Drake's originally reported value for the electron
affinity of Hydrogen is precisely the value reported as ``exact'' in table(%
\ref{EAtab}), however, this is only coincidental. Also, the result from
Pekeris in table(\ref{EAtab}) includes only first order mass polarization
correction and Drakes relativistic corrections $\Delta E_{corr}.$ 

%TCIMACRO{\TeXButton{B}{\begin{table}[tbp] \centering}}
%BeginExpansion
\begin{table}[tbp] \centering%
%EndExpansion
\begin{tabular}{llll}
\hline\hline
& Hydrogen & Deuterium & Tritium \\ \hline
$E_H-E_{H^{-}}$ & 6083.4058$cm^{-1}$ & 6087.0201$cm^{-1}$ & 6088.2233$cm^{-1}
$ \\ 
$\Delta E_{corr}$ & .307505$cm^{-1}$ & .307589$cm^{-1}$ & .307616$cm^{-1}$
\\ 
$EA$ & 6083.0983$cm^{-1}$ & 6086.7126$cm^{-1}$ & 6087.9157$cm^{-1}$ \\ 
``Exact'' estimate & 6083.0994$cm^{-1}$ & 6086.7137$cm^{-1}$ & 6087.9168$%
cm^{-1}$ \\ \hline
Drake\cite{Drake88}(corrected) & 6083.097278$cm^{-1}$ &  &  \\ 
Pekeris\cite{Pekeris62}(using $\Delta E_{corr}$) & 6083.0909$cm^{-1}$ &  & 
\\ 
Lykke\cite{Lykke91}(experiment) & 6082.99$\pm 0.15cm^{-1}$ & 6086.2$\pm
0.6cm^{-1}$ &  \\ \hline\hline
\end{tabular}
\caption{Electron affinity of Hydrogen, Deuterium and Tritium. The  term $ \Delta E_{corr} $
contains relativistic, relativistic recoil, Lamb shift and finite nuclear size corrections   \label{EAtab}}%
%TCIMACRO{\TeXButton{E}{\end{table}}}
%BeginExpansion
\end{table}%
%EndExpansion

\subsection{Conclusion}

This work has provided reliable values for the electron affinity of
Hydrogen, Deuterium and Tritium. These electron affinity values are the
consequence of improved rigorous variational energy upper bounds for these
systems. Constructing wave functions using a non-adiabatic Hamiltonian has
provided results unattainable with the infinite nuclear mass approximation
plus Rydberg scaling and exact mass polarization correction. It was shown
that 9 nano-hartrees of the energy shift due to nuclear motion cannot be
accounted for by these corrections. Thus is appears that the only way to
improve on our computed energy bounds is by rigorous non-adiabatic energy
calculations, perhaps using one of the outstanding basis sets used in very
accurate infinite nuclear mass calculations. These results may be important
for researchers testing the validity of QED calculations such as the Lamb
shift and other small energy corrections and may also be significant for
researchers attempting to compute values for fundamental physical constants.
We hope that others will confirm and improve on our results, and that
experimentalists will push the envelope of error bars closer to the accuracy
of the theoretical calculations. 

\noindent \textbf{Acknowledgment}

This work was supported by the National Science Foundation with grant
CHE-9300497

\bibliographystyle{aip}
\bibliography{4-prefs,hmhe,mcalc}

\end{document}
